---
title: "Rlab: continuous spatial index"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```


```{r load_packages, include=FALSE}

library(sp)
library(ggmap)
library(tidyverse)
library(magrittr)
library(gstat)

```


```{r load_data, echo=FALSE}

setwd(here::here())

load("data/city_boundary.RData") #loads sd_city variable

water_testing <- read.csv("data/indicator_bacteria_tests_datasd_v1.csv")

sites_ll <- read.csv("data/sample_sites_datasd_v1.csv")

load("data/sd_tracts_income.RData")

load("data/sd_grid.RData")

# prepare data 

coordinates(sites_ll) <- c("lng", "lat")
proj4string(sites_ll) <- CRS("+proj=longlat")
sites <- spTransform(sites_ll, proj4string(sd_city))

```

### I. Gathering and exploring the data

1. [2 pts] What type of R object is `sd_city`? What coordinate reference system is used?

- `sd_city` is `r class(sd_city)[[1]]`

- Coordinate reference system:

```{r}

proj4string(sd_city)

```


2. [1 pts] What does line 7 do?

- Transform for map projection and datum transformation. `sites_11` is the object to be transfer. And `proj4string(sd_city)` is the coordinate references CRS object.


3. [4 pts] Make a plot that shows: (i) the locations of the measurement sites, (ii) the boundary of the city, and (iii) basic contextual features of San Diego, such as major roadways.

```{r fig.height=4}

sites <- spTransform(sites, CRS("+proj=longlat"))

map_stamen <- get_stamenmap(bbox = c(left = -117.3, bottom = 32.55,
                                     right = -116.9, top = 33.05))

ggmap(map_stamen) +
  geom_point(aes(x = lng, y = lat), data = as.data.frame(sites))

```


```{r}

sites$cl2_total <- sapply(as.character(sites_ll$site_ID), function(site){
  median(water_testing$cl2_total[water_testing$source == site], na.rm = T)
})

```


4. [2 pts] There are 9 NAs in `sites$cl2_total`. Why?

- Because there're 9 sites that exist in `sites_ll$site_ID` but not `water_testing$source`


5. [2 pts] Write your own code that creates a new variable called `sites_sub` that is just like `sites`, but has the 9 sample sites with `NA` values for `cl2_total` removed.

```{r echo=TRUE}

sites_sub <- sites

sites_sub@data %<>% filter(!is.na(cl2_total))

```


6. [1 pts] What is the coordinate reference system for the tract data?

```{r}

sd_tracts@proj4string

```


7. [1 pts] What is the name of the census tract with the lowest median household income? Highest?

- Name of census tract with lowest median income: `r sd_tracts@data$NAME[sd_tracts@data$median_income == min(sd_tracts@data$median_income)]`

- Name of census tract with highest median income: `r sd_tracts@data$NAME[sd_tracts@data$median_income == max(sd_tracts@data$median_income)]`


8. [4 pts] Provide executable code that adds a new column to `sites_sub` called `median_income` with the appropriate values of median household income. Check that your first few values match the ones below.

```{r}
# 
sd_tracts %<>% spTransform(CRS("+proj=longlat"))

SpatialPointsDataFrame()
# 
# sites_sub_tract <- over(SpatialPolygons(sd_tracts@polygons), SpatialPoints(sites_sub@coords))
# 
# sites_sub_tract_table <- table(sites_sub_tract)
# 
# x <- rep(0, length(sd_tracts))
# 
# x[as.numeric(names(sites_sub_tract_table))] <- sites_sub_tract_table
# 
# sd_tracts$x <- x
# 
# spplot(sd_tracts, zcol = "x", xlim = c(-117.3, -116.8), ylim = c(32.5, 33))

# set.seed(1331)
# pts = cbind(1:5, 1:5)
# dimnames(pts)[[1]] = letters[1:5]
# df = data.frame(a = 1:5)
# row.names(df) = letters[5:1]
# 
# SpatialPointsDataFrame(pts, df)
# 
# 
# sites_sub@data
# 
# sites_sub@coords
# 
# 
# SpatialPointsDataFrame(as.data.frame(sites_sub@data), sd_tracts@data$median_income)

```


9. [4 pts] Fit a non-spatial linear model using `lm()` with median household income as the sole predictor of chlorine levels. Give a point estimate and 95% confidence interval for the effect of median household income on chlorine levels. How does the estimate compare to your expectations?


### II. Spatial dependence + interpolation





12. [3 pts] Use the krige function to perform universal kriging of chlorine levels for all locations in `sd_grid`, taking into account the effect of household income. Use whichever variogram model you think fit best. Provide the code you used to make your predictions, and report estimates of parameters in the variogram model.

```{r}

```


13. [3 pts] Make a map of the predictions, again with contextual information included (hint: if you use ggmap(), you’ll need to transform to longitude/latitude coordinates.)

14. [3 pts] Below is a map I made based on my model. How does it compare to yours? What does your map show that mine doesn’t?

15. [2 pts] If you look closely, you can see some interesting shapes showing up in my figure. What do they correspond to? Why do you think they’re visible in the predicted surface?















